Post-Disturbance Diversity Trajectories in a Neotropical Forest

### 2018-02-06 Manuscript1.Rmd
Manuscript update

### Study site 
Paracou station in French Guiana (5°18'N and 52°53'W) is located in a lowland tropical rain forest corresponding to a tropical wet climate with mean annual precipitation averaging 2980 mm.y^-1^ (30-y period) with a 3-months dry season (< 100 mm.months^-1^) from mid-August to mid-November, and a one-month dry season in March @Wagner2011. 
Elevation ranges between 5 and 50 m and mean annual temperature is 26°C. 
Soils correspond to thin acrisols over a layer of transformed saprolite with low permeability generating lateral drainage during heavy rains.
The experiment corresponds to a network of twelve 6.25ha plots where three logging, thinning and fuelwood treatments on an intensity gradient.
Disturbance treatments were attributed according to a randomized plot design with three replicate blocks of four plots. 
The logging correspond to averages of 10 trees removed per hectare with a diameter at 1.3 m height (DBH) above 50 cm for treatment 1 (T1), 32 trees/ha above 40 cm DBH for treatment 2 (T2) and 40 trees above 40 cm DBH for treatment 3 (T3). 
Treatments T2 and T3 included besides logging of 10trees/ha the thinning by poison girdling \ref{@Blanc2009}.

Table 1: Intervention table, summary of the disturbance intensity for the 4 plot treatments in Paracou.

----------------------------------------------------------------------------------------------------------------------
                      ***Timber***            ***Thinning***               ***Fuelwood***               *%AGB lost*
-------------     --------------------     -------------------------     ------------------------     ---------------
***Control***                                                                                              0

***T1***            DBH ≥ 50 cm,
                    commercial species,                                                                [12%-33%]
                    ~ 10 trees/ha

***T2***            DBH ≥ 50 cm,            DBH ≥ 40 cm,
                    commercial species,    non-valuable species,                                        [33%-56%]
                    ~ 10 trees/ha          ~ 30 trees/ha

***T2***            DBH ≥ 50 cm,            DBH ≥ 50 cm,                   40 cm ≤ DBH ≤ 50 cm
                    commercial species,    non-valuable species,           non-valuable species,        [33%-56%]
                    ~ 10 trees/ha          ~ 15 trees/ha                   ~ 15 trees/ha
----------------------------------------------------------------------------------------------------------------------

@@

# Discussion
-	The disturbance significantly impact the functional and taxonomic composition and diversity of the stands for more than 30 years 
-	Important differences among plots of same treatments: possible impact of initial diversity composition, local abiotic parameters (topography, soils, surrounding stands).   Impacts depend on the disturbance intensity, but further parameters may be involved. 
-	The shift in functional composition differs from the null model: biotic and abiotic filters are at stake but only tangible from 10/15 years after logging. Discussion with Cannon, 1998: they found no different process than rarefaction for 8 years after logging, but the time might have been too short-> divergence due to recruitment and not rarefaction after random species disappearance. Same conclusions otherwise.
-	Disturbance involves new filters for species settling in the stands, probably according to their functional strategy. Disturbance generates a functional shift of communities toward acquisitive strategies. Ecosystems function rely on abundant species: changes in stands functioning is then expected.  (Flores, 2006  for differential sapling success after disturbance).
-	If the recovery is not achieved 30 years after disturbance, it is on the way to. Only one plot displayed a still increasing Simpson diversity, and all plots showed a stabilization or a decrease of the composition distancing from 1989 inventories.
-	To go further we should focus on recruitment which would also highlight the mechanisms underlying these shifts and dynamics and explicit the environmental filter.

>>>>

### Study site 
Paracou station in French Guiana (5°18'N and 52°53'W) is located in a lowland tropical rain forest corresponding to a tropical wet climate with mean annual precipitation averaging 2980 mm.y^-1^ (30-y period) with a 3-months dry season (< 100 mm.months^-1^) from mid-August to mid-November, and a one-month dry season in March [@Wagner2011]. 
Elevation ranges between 5 and 50 m and mean annual temperature is 26°C. 
Soils correspond to thin acrisols over a layer of transformed saprolite with low permeability generating lateral drainage during heavy rains.
The experiment corresponds to a network of twelve 6.25ha plots that have undergone a gradient of three logging, thinning and fuelwood treatments.
Disturbance treatments were attributed according to a randomized plot design with three replicate blocks of four plots. 
The disturbance corresponds to averages of 10 trees removed per hectare with a diameter at 1.3 m height (DBH) above 50 cm for treatment 1 (T1), 32 trees/ha above 40 cm DBH for treatment 2 (T2) and 40 trees above 40 cm DBH for treatment 3 (T3). 
Treatments T2 and T3 besides included the thinning of trees by poison girdling [@Blanc2009].

Table 1: Intervention table, summary of the disturbance intensity for the 4 plot treatments in Paracou.
---------------------------------------------------------------------------------------------------------------------------------
                          ***Timber***                 ***Thinning***              ***Fuelwood***           **%AGB lost**

-----------------    ------------------------    ------------------------    ------------------------    ------------------------
    ***Control***                                                                                                   0

-----------------    ------------------------    ------------------------    ------------------------    ------------------------
         ***T1***     DBH ≥ 50 cm,                                                                                               
                      commercial species,                                                                      [12%-33%]         
                      ~ 10 trees/ha

-----------------    ------------------------    ------------------------    ------------------------    ------------------------
         ***T2***     DBH ≥ 50 cm,                DBH ≥ 40 cm,                                                                   
                      commercial species,         non-valuable species,                                        [33%-56%]         
                      ~ 10 trees/ha               ~ 30 trees/ha

-----------------    ------------------------    ------------------------    ------------------------    ------------------------
         ***T3***     DBH ≥ 50 cm,                DBH ≥ 50 cm,                 40 cm ≤ DBH ≤ 50 cm                               
                      commercial species,         non-valuable species,        non-valuable species,           [33%-56%]         
                      ~ 10 trees/ha               ~ 15 trees/ha                ~ 15 trees/ha

---------------------------------------------------------------------------------------------------------------------------------

@@

# Discussion
-	The disturbance significantly impact the functional and taxonomic composition and diversity of the stands for more than 30 years 
-	Important differences among plots of same treatments: possible impact of initial diversity composition, local abiotic parameters (topography, soils, surrounding stands).   Impacts depend on the disturbance intensity, but further parameters may be involved. 
-	The shift in functional composition differs from the null model: biotic and abiotic filters are at stake but only tangible from 10/15 years after logging. Discussion with Cannon, 1998: they found no different process than rarefaction for 8 years after logging, but the time might have been too short-> divergence due to recruitment and not rarefaction after random species disappearance. Same conclusions otherwise.
-	Disturbance involves new filters for species settling in the stands, probably according to their functional strategy. Disturbance generates a functional shift of communities toward acquisitive strategies. Ecosystems function rely on abundant species: changes in stands functioning is then expected.  (Flores, 2006  for differential sapling success after disturbance).
-	If the recovery is not achieved 30 years after disturbance, it is on the way to. Only one plot displayed a still increasing Simpson diversity, and all plots showed a stabilization or a decrease of the composition distancing from 1989 inventories. however this is the impact of a single disturbance event: the repetition may have different impact (@Olander2005 demonstrated the impact of logging on soils nutrient and compaction)
-	To go further we should focus on recruitment which would also highlight the mechanisms underlying these shifts and dynamics and explicit the environmental filter.
